We are not the 99 percent: quantifying asphericity in the distribution of Local Group satellites

### 2017-12-03 paper/paper.tex
abstract

\begin{abstract}
We quantify the asphericity in the spatial distribution of the
brightest satellites around the Milky Way (MW) and M31, the dominant
galaxies in the Local Group (LG).
We use this approach on simulations to build an explicit probability
distribution of finding an aspherical  satellite  distribution around
galaxies similar to the Milky Way and M31. 
This probability distribution allows us to estimate the atipicallity
of the satellite distributions in the LG even when the underlying
simulations do not have enough systems that resemble the LG satellite
anisotropy. 
This methodology is straightforward to implement, allowing
for an straighforward quantification of aspherical satellite
distributions both in simulations and observations.
We demonstrate the method using three different simulations:
Illustris-1,  Illustris-1-DarkMatter and ELVIS simulations. 
We find as a common trend that at most of $0.20\%$ the pairs are
expected  to have satellite distributions with the same degree of
deviation from sphericity as the Local Group.
The three simulations show that the satellite distribution in M31 is
common while the very planar MW satellite distribution is hard to
find.
We show that in these simulations between $19\%$ to $24\%$ of the pairs
have a halo with a satellite distribution similar to M31, while only
$0.80\%$, at most, have a satellite distribution as planar as the MW.
This characteristic places the LG as at the level of a 3-$\sigma$
outlier in context provided by LCDM simulations.
\end{abstract}

>>>>

\begin{abstract}
We quantify the asphericity in the spatial distribution of the
brightest satellites around the Milky Way (MW) and M31, the dominant
galaxies in the Local Group (LG).
We use the same approach on simulations to build an explicit probability
distribution for this asphericity around galaxies similar to the MW and M31.   
This allows us to estimate the atipicallity
of the satellite distributions in the LG even when the underlying
simulations do not have enough systems that fully resemble the LG.
We demonstrate the method using three different simulations:
Illustris-1,  Illustris-1-DarkMatter and ELVIS. 
We find as a common trend that at most of $0.2\%$ the pairs are
expected  to have satellite distributions with the same degree of
deviation from sphericity as the Local Group; between $19\%$ to $24\%$
of the pairs have a halo with a satellite distribution as aspherical
as in M31, while only $0.80\%$, at most, have a satellite distribution
as planar as in the MW. These results place the LG at the level of a
3-$\sigma$ outlier in context provided by LCDM simulations. 
\end{abstract}
